---
title: "Part II: Teaching Practice (<= 30 min)"
subtitle: "Sampling Distribution of the Mean"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r, echo = FALSE, warning = FALSE, message = FALSE}
library(tidyverse)
knitr::opts_chunk$set(echo = FALSE, fig.width = 8, fig.height = 2, 
					  fig.align = "center", warning = FALSE)
```

## Setting

I'll be demonstrating the __simulation__ method for student engagement.

Prerequisite knowledge that students would be expected to know:

- Sampling from a population (distribution) in R.
- Multi-step simulations.
- How to interpret and calculate mean and quantiles, both empirical and true.
- Conceptually build and read a histogram, but not necessarily know how to make one.
- The concept of a sample/empirical mean vs. the true mean, and the LLN.

## Learning Objectives

From today's lesson, students are expected to be able to:

- Define a sampling distribution, particularly pointing out the difference between a population distribution.
- Generate a sample from a sampling distribution of the mean, given a population distribution.

**Hint**: Use these learning objectives when studying for your quiz!

## Overview

So far, we've learned about *population distributions*, and how they can be used to explain the uncertainty of an unknown outcome. Here are the population distributions of three examples we've been working with:

```{r, echo = FALSE, warning = FALSE, message = FALSE}
dir("supplementary", full.names = TRUE) %>% 
	walk(source)
cowplot::plot_grid(
  tibble(x = expense$qdist(c(0, 0.99))) %>%
    ggplot(aes(x)) +
    stat_function(fun = expense$ddist) +
    theme_bw() +
    # ylab("Density") +
  	scale_x_continuous(
  		"Monthly Expense", 
  		labels = scales::dollar_format(scale = 1/1000, 
  									   suffix = "K", 
  									   accuracy = 0.1)
  	) +
  	scale_y_continuous("Density"),
  tibble(x = c(-0.5, 1.5)) %>% 
    ggplot(aes(x)) +
    stat_function(fun = octane$ddist) +
    theme_bw() +
    labs(x = "Octane Purity",
         y = "Density"),
  ggplot(los$pmf, aes(ndays, p)) + 
  	geom_col() +
  	ylab("Probability") +
  	xlab("Length of Stay") +
  	theme_bw(),
  nrow = 1
)
```

We saw that there are two "versions" of the mean:

- The __empirical mean__ is calculated using data, and is an average: $1/n \sum_i X_i$
- The __true mean__ is what the empirical mean converges to as we collect more and more data. It can be calculated from a population distribution's density as $\int_x x f(x) dx.$

Hence, the empirical mean is an _estimate_ of the true mean.

__Today__: If the empirical mean is an _estimate_ of the true mean, how can we communicate _how good_ of an estimate it is? The answer is, by using the distribution of the empirical mean! This distribution has a special name: it's called a _sampling distribution_. 

## Concept 1

**A combination of random variables is also a random variable.**

For example, if $X$, $Y$, and $Z$ are random variables, then...

- $X^Y$ is also a random variable.
- $\exp(X) + Z\sin(Y)$ is also a random variable.
- ...

Remember, even though these quantities are made up of many random variables, they still calculate to a single number -- hence why they amount to a new single random variable.

In fact, we've seen a similar concept in the context of [multi-step simulations](https://ubc-mds.github.io/DSCI_551_stat-prob-dsci/lectures/simulation.html#multi-step-simulations-10-min), where we calculated total "gang demand" from ships arriving at port, and each making a random demand.

## Concept 2

**The empirical mean has a distribution, and is called a _sampling distribution_**.

For example, consider the mean of 3 observations of **Length of Stay**. We'll get different means depending on the sample that's drawn:

| Ship | Lengths of Stay | Mean |
|------|-----------------|------|
| Ships 1 - 3 | 1, 3, 3 | 2.333 |
| Ships 4 - 6 | 2, 1, 3 | 2 |
| Ships 7 - 9 | 2, 2, 4 | 2.667 |

This is because the underlying sample is random.

To formalize concepts, let's call three random draws $X_1, X_2, X_3$. Then their mean is
$$\bar{X} = \frac{X_1 + X_2 + X_3}{3},$$ 
which is a random variable (because each of $X_1, X_2, X_3$ are). This random variable's distribution is called a _sampling distribution_.

## Demonstration

Let's use simulation to view the sampling distribution of 10 **Length of Stay** observations.

Steps:

1. Everyone generate their own 10 observations from the **Length of Stay** distribution, and calculate the mean.
	- To start you off, here is code to generate 10 observations:

```
sample(1:5, size = 10, replace = TRUE, prob = c(0.25, 0.35, 0.2, 0.1, 0.1))
```

2. Submit your mean to [this google form](https://docs.google.com/forms/d/e/1FAIpQLSdnQXaFOjKM_Wzsf0O9tMSSjIpXA0Nk34JTUjqtaFP5hy9N0Q/viewform?usp=sf_link).
	- [Spreadsheet of results](https://docs.google.com/spreadsheets/d/1mKuprd-g-IrPekcVuVSkax0_4Kty0NznjFnvAETN32Y/edit#gid=1852227229)


## Other Examples

Here are the sampling distributions of the mean of 10 observations for the three examples we're considering. They are obtained by generating 1000 means each.

```{r, echo = FALSE}
# N = number of iterations, n = sample size, rdist = function that
#  generates a sample.
set.seed(1)
get_means <- function(N, n, rdist) {
	map_dbl(1:N, function(x) mean(rdist(n)))
}
tribble(
	~ example, ~ rdist,
	"Expense", expense$rdist,
	"Octane", octane$rdist,
	"Length of Stay", los$rdist
) %>% 
	transmute(example = as_factor(example),
			  means = map(rdist, ~ get_means(1000, n = 10, rdist = .x))) %>% 
	unnest(means) %>% 
	ggplot(aes(means)) +
	facet_wrap(~ example, scales = "free") +
	geom_histogram(aes(y = ..density..), bins = 25) +
	labs(x = "Empirical Mean",
		 y = "Density") +
	theme_bw()
```
